---
title: "Data Analysis Project Proposal"
author: "STAT 420, Summer 2021, Michael McClanahan (NetID: mjm31) and Jake Goodman (NetID: jakeg5)"
date: ''
output:
  html_document: 
    theme: readable
    toc: yes
  pdf_document: default
urlcolor: cyan
---

***

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
options(scipen = 1, digits = 4, width = 80, fig.align = "center")
```

### Contributing Students

The following students will be contributing to the group project:

- Michael McClanahan (NetID: mjm31)
- Jake Goodman (NetID: jakeg5)


### Tentative Project Title

Direct Marketing Charitable Contribution Prediction

### Data File Description

- General Background 
  - This data file includes information on donors who have made charitable contributions to a National Veteran's Agency.
- Dataset Characteristics
  - 7600 records
  - 19 variables
    - 17 numeric variables
    - 2 categorical variables
- Variables of Interest
  - Response Variable
    - `Average.Gift`: will be used as the response variable for this project to understand the factors behind the average donation amount someone makes
  - Explanatory Variables
    - `Age`: the age of the donor
    - `Own.Home.`: whether or not someone owns a home or if it's unknown
    - `Num.Children`: the number of children, if any, that the donor has
    - `Total.Wealth`: a wealth rating from 0-9 with 9 being the wealthiest that is based on median family income and population statistics
    - `Sex`: the gender of the donor
    - `Number.of.Gifts`: the number of donations made
    - `Time.Between.Gifts`: the avg number of months between donations
    - `Largest.Gift`: the largest donation made (in dollars)
    - `Smallest.Gift`: the smallest donation made (in dollars)
    - `Other.Gifts`: number of times the donor has responded to a mail order donation offer

### Data Set Background Information

### Statement of Interest

Being that there are over a million charities in the U.S. (https://nccs.urban.org/publication/nonprofit-sector-brief-2019#finances), there is reliance on charitable contributions to ensure they're able to continue to deliver their services and value to their respective communities. To best be able to do that, a better understanding of the ideal donor would be beneficial to hopefully maximize the amount of donations they receive. With this project, we can hopefully develop a good understanding of what characteristics an ideal donor might have.

### Evidence that the data can be loaded into R

```{r}
nvo_data = read.table('nvo.txt', sep = '\t',header = TRUE, fileEncoding = "UTF-8")
head(nvo_data)
```

***
      
